Monte Carlo and bootstrap for all parameters with a grid

### 2024-08-01 Boot_grid_ter.Rmd
A third way to use the grid

---
title: "Monte Carlo and bootstrap for all parameters with a grid"
output: pdf_document
date: "2024-07-26"
---

@@

The goal of this document is to create a large number of Monte Carlo iterations to produce estimators (for the number of trees, the volume of trees and the ratio volume of trees / number of trees), variances of these estimators and confidence intervals. The same elements are also computed for each bootstrap series derived from a smaller number of Monte Carlo iterations. By comparing the two results, we may verify the invalidity of the bootstrap strategy to find the estimation of the variances.

@@

for (l in 0:((1000/c_dim)-1)) {
  if (l==0){
    x_init <- i*c_dim
  }else if ((3*l+i)*c_dim < (10)*c_dim){
    x_init <- (3*l+i)*c_dim
  }else{
    x_init <- (((3*l+i)*c_dim)%%(c_dim*10))
  }
  k <- 0
  while (x_init + 10*c_dim*k < 1000) {
    a <- runif(1, x_init + 10*c_dim*k, x_init + 10*c_dim*k + c_dim)
    b <- runif(1, l*c_dim, (l+1)*c_dim)
    random_points <- rbind(random_points, list(a,b))
    k <- k+1
  }
}
colnames(random_points) <- list("x", "y")
return(random_points)
}
```

@@

# Monte Carlo alone

10 000 Monte Carlo iterations are appropriated.

```{r}
### Function to automate Monte Carlo method

MonteCarlo <- function(n, B){
  # B: number of iterations for MC

  p <- sum(rep((1/1000)*(1/1000), n)) # inclusion density
  MC_df <- data.frame(useless = rep(0, n))
  MC_est <- data.frame(useless = rep(0, 3), row.names = c("est_nb", "est_vol", "est_ratio"))

  for (b in 1:B) { # All in one "for" loop
    # Sampling over the whole territory
    df_grid <- sampling_grid(i,c_dim)
    a1 <- df_grid$x
    a2 <- df_grid$y
    # n <- nrow(df_grid) # sample size
    # p <- sum(rep((1/1000)*(1/1000), n)) # inclusion density
    # Variables of interest derived from the sample
    var <- mapply(circle_all, a1, a2)
    MC_b <- data.frame(x = a1, y = a2, z = t(var)[,1], t = t(var)[,2])
    MC_df <- cbind(MC_df, MC_b)
    colnames(MC_df)[(4*(b-1)+2):(4*(b-1)+2+3)] <- c(paste("x", b, sep = ""), paste("y", b, sep = ""), paste("nb", b, sep = ""), paste("vol", b, sep = ""))
    est_b <- t(data.frame(x = sum(MC_df[4*(b-1)+2+2] /p), y = sum(MC_df[4*(b-1)+2+3] /p), z = (sum(MC_df[4*(b-1)+2+3] /p))/(sum(MC_df[4*(b-1)+2+2]/p))))
    MC_est <- cbind(MC_est, est_b)
    colnames(MC_est)[b+1] <- paste("ech", b, sep = "") 
  }
  MC_df$useless <- NULL
  MC_est$useless <- NULL
  MC_var <- c(var(as.vector(t(MC_est[1,]))), var(as.vector(t(MC_est[2,]))), var(as.vector(t(MC_est[3,]))))

  # Confidence intervals
  norm_CI <- data.frame(useless = rep(0, 2))
  perc_CI <- data.frame(useless = rep(0, 2))
  rev_perc_CI <- data.frame(useless = rep(0, 2))
  for (l in 1:3) { # Loop: one estimator each time
    est_init <- as.vector(t(MC_est[l,]))[1]
    sorted_est <- sort(as.vector(t(MC_est[l,])))
    lim_inf <- sorted_est[floor(B*2.5/100)]
    lim_sup <- sorted_est[B - floor(B*2.5/100)]
    perc_CI <- cbind(perc_CI, data.frame(c(lim_inf, lim_sup)))
    rev_perc_CI <- cbind(rev_perc_CI, data.frame(c(est_init*2 - lim_sup, est_init*2 - lim_inf)))
    norm_CI <- cbind(norm_CI, data.frame(c(est_init-qnorm(0.975)*sqrt(MC_var[l]), est_init+qnorm(0.975)*sqrt(MC_var[l]))))
  }
  norm_CI$useless <- NULL
  perc_CI$useless <- NULL
  rev_perc_CI$useless <- NULL
  colnames(norm_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
  colnames(perc_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
  colnames(rev_perc_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
  all_CI <- list(Normal = norm_CI, Percentile = perc_CI, `Reverse percentile` = rev_perc_CI)

  return(list(`Local Variables` = MC_df, Estimations = MC_est, Variances = MC_var, `Confidence Intervals` = all_CI))
  # Local Variables: local variables (number and volume) for each point in each sample
  # Estimations: 3 Estimations (number, volume, ratio volume / number) for each sample
  # Variances: Monte Carlo variance for each of the totals / ratio
  # Confidence Intervals: 3 different forms of 95% confidence intervals, for each total / ratio
}
```

@@

  p <- sum(rep((1/1000)*(1/1000), n))
  BOOT_df <- list() # All local variables
  BOOT_est <- list() # All estimations
  BOOT_var <- list() # Variances
  BOOT_CI <- list() # Confidence Intervals
  for (b in 1:B1) { # One MC iteration at a time
    # Uniform sampling over the whole territory
    df_grid <- sampling_grid(i,c_dim)
    a1 <- df_grid$x
    a2 <- df_grid$y
    # n <- nrow(df_grid) # sample size
    # p <- sum(rep((1/1000)*(1/1000), n)) # inclusion density
    var <- mapply(circle_all, a1, a2)
    df_b <- data.frame(x = a1, y = a2, z = t(var)[,1], t = t(var)[,2])
    colnames(df_b) <- c(paste("x.MC", b, sep = ""), paste("y.MC", b, sep = ""), paste("nb.MC", b, sep = ""), paste("vol.MC", b, sep = ""))
    est_b <- t(data.frame(x = sum(df_b[3] /p), y = sum(df_b[4] /p), z = (sum(df_b[4] /p))/(sum(df_b[3]/p))))

@@

## Monte Carlo alone

res1 <- MonteCarlo(n, 10000)
# Just one circle to begin with

```{r}
cat("Monte Carlo means of estimations: \n - for the number of trees:", mean(as.vector(t(res1$Estimations[1,]))), "\n - for the volume of trees:", mean(as.vector(t(res1$Estimations[2,]))), "\n - for the ratio volume / number of trees:", mean(as.vector(t(res1$Estimations[3,]))))

# Example:
## Monte Carlo means of estimations:
##  - for the number of trees: 29529.27
##  - for the volume of trees: 5428.676
##  - for the ratio volume / number of trees: 0.1838376
```

```{r}
cat("Monte Carlo variances: \n - for the number of trees:", res1$Variances[1], "\n - for the volume of trees:", res1$Variances[2], "\n - for the ratio volume / number of trees:", res1$Variances[3])

# Example:
## Monte Carlo variances:
##  - for the number of trees: 263478.6
##  - for the volume of trees: 15475.27
##  - for the ratio volume / number of trees: 6.470082e-06
```

```{r}
cat("95% confidence intervals using the normal method:\n - for the number of trees: [", res1$`Confidence Intervals`$Normal[1,1], ";", res1$`Confidence Intervals`$Normal[2,1], 
    "]\n - for the volume of trees: [", res1$`Confidence Intervals`$Normal[1,2], ";", res1$`Confidence Intervals`$Normal[2,2], 
    "]\n - for the ratio volume / number of trees: [", res1$`Confidence Intervals`$Normal[1,3], ";", res1$`Confidence Intervals`$Normal[2,3],
    "]")

# Example:
## 95% confidence intervals using the normal method:
##  - for the number of trees: [ 28787.75 ; 30799.86 ]
##  - for the volume of trees: [ 5115.539 ; 5603.177 ]
##  - for the ratio volume / number of trees: [ 0.1748962 ; 0.1848671 ]
```

```{r}
cat("95% confidence intervals using the percentile method:\n - for the number of trees: [", res1$`Confidence Intervals`$Percentile[1,1], ";", res1$`Confidence Intervals`$Percentile[2,1], 
    "]\n - for the volume of trees: [", res1$`Confidence Intervals`$Percentile[1,2], ";", res1$`Confidence Intervals`$Percentile[2,2], 
    "]\n - for the ratio volume / number of trees: [", res1$`Confidence Intervals`$Percentile[1,3], ";", res1$`Confidence Intervals`$Percentile[2,3],
    "]")

# Example:
## 95% confidence intervals using the percentile method:
##  - for the number of trees: [ 28533.3 ; 30557.75 ]
##  - for the volume of trees: [ 5188.069 ; 5673.772 ]
##  - for the ratio volume / number of trees: [ 0.1787598 ; 0.1888164 ]
```

```{r}
cat("95% confidence intervals using the reverse percentile method:\n - for the number of trees: [", res1$`Confidence Intervals`$`Reverse percentile`[1,1], ";", res1$`Confidence Intervals`$`Reverse percentile`[2,1], 
    "]\n - for the volume of trees: [", res1$`Confidence Intervals`$`Reverse percentile`[1,2], ";", res1$`Confidence Intervals`$`Reverse percentile`[2,2], 
    "]\n - for the ratio volume / number of trees: [", res1$`Confidence Intervals`$`Reverse percentile`[1,3], ";", res1$`Confidence Intervals`$`Reverse percentile`[2,3],
    "]"
    )

# Example:
## 95% confidence intervals using the reverse percentile method:
##  - for the number of trees: [ 29029.86 ; 31054.31 ]
##  - for the volume of trees: [ 5044.944 ; 5530.647 ]
##  - for the ratio volume / number of trees: [ 0.1709469 ; 0.1810035 ]
```

@@

```{r}
cat("Mean of the means of estimations obtained through each bootstrap process:\n - for the number of trees:", mean(EST_nb), "\n - for the volume of trees:", mean(EST_vol), "\n - for the ratio volume / number of trees:", mean(EST_ratio))

# Example:
## Mean of the means of estimations obtained through each bootstrap process:
##  - for the number of trees: 29546.03
##  - for the volume of trees: 5431.112
##  - for the ratio volume / number of trees: 0.1838134
```

@@

```{r}
cat("Mean of the variance of bootstrap samples:\n - for the number of trees:", mean(VAR_nb), "\n - for the volume of trees:", mean(VAR_vol), "\n - for the ratio volume / number of trees:", mean(VAR_ratio))

# Example:
## Mean of the variance of bootstrap samples:
##  - for the number of trees: 2630037
##  - for the volume of trees: 99634.96
##  - for the ratio volume / number of trees: 1.033748e-05
```

@@

```{r}
diff_percent <- function(x,y)100*(max(x,y)-min(x,y))/min(x,y)
cat("Difference between the two estimations of variance: \n - for the number of trees: ", diff_percent(res1$Variances[1], mean(VAR_nb)), "%", "\n - for the volume of trees: ", diff_percent(res1$Variances[2], mean(VAR_vol)), "%\n - for the ratio volume / number of trees: ", diff_percent(res1$Variances[3], mean(VAR_ratio)), "%", sep = "")

# Example:
## Difference between the two estimations of variance:
##  - for the number of trees: 898.1975%
##  - for the volume of trees: 543.8335%
##  - for the ratio volume / number of trees: 59.77348%
```

>>>>

---
title: "Monte Carlo and bootstrap for all parameters with a grid"
output: pdf_document
date: "2024-08-01"
---

@@

Motivation: specific sampling from a grid, random small square inside a first bunch of 10 small squares, for a beginning. Then, systematic sampling without any shifts when changing line.

The goal of this document is to create a large number of Monte Carlo iterations to produce estimators (for the number of trees, the volume of trees and the ratio volume of trees / number of trees) and variances of these estimators. The same elements, along with confidence intervals, are also computed for each bootstrap series derived from a smaller number of Monte Carlo iterations. By comparing the two results, we may verify the invalidity of the bootstrap strategy to find the estimation of the variances.
We reuse the results from another file for the function that exploits Monte Carlo alone.

@@

for (l in 0:((1000/c_dim)-1)) {
  x_init <- i*c_dim
  k <- 0
  while (x_init + 10*c_dim*k < 1000) {
    a <- runif(1, x_init + 10*c_dim*k, x_init + 10*c_dim*k + c_dim)
    b <- runif(1, l*c_dim, (l+1)*c_dim)
    random_points <- rbind(random_points, list(a,b))
    k <- k+1
  }
}
colnames(random_points) <- list("x", "y")
return(random_points)
}
```

```{r}
# Small example
ex_grid <- sampling_grid(i, c_dim)
```

```{r}
plot(ex_grid, xlab="", ylab="", main = "Random points within the grid", las = 1)
```

```{r}
plot(ex_grid, xlab="", ylab="", main = "Only one point", xlim = c(0, 100), ylim = c(0, 10), las = 1)
```

@@

@@

  p <- sum(rep((1/1000)*(1/1000), n))
  BOOT_df <- list() # All local variables
  BOOT_est <- list() # All estimations
  BOOT_var <- list() # Variances
  BOOT_CI <- list() # Confidence Intervals
  for (b in 1:B1) { # One MC iteration at a time
    # Uniform sampling over the whole territory
    df_grid <- sampling_grid(i, c_dim)
    a1 <- df_grid$x
    a2 <- df_grid$y
    # n <- nrow(df_grid) # sample size
    # p <- sum(rep((1/1000)*(1/1000), n)) # inclusion density
    var <- mapply(circle_all, a1, a2)
    df_b <- data.frame(x = a1, y = a2, z = t(var)[,1], t = t(var)[,2])
    colnames(df_b) <- c(paste("x.MC", b, sep = ""), paste("y.MC", b, sep = ""), paste("nb.MC", b, sep = ""), paste("vol.MC", b, sep = ""))
    est_b <- t(data.frame(x = sum(df_b[3] /p), y = sum(df_b[4] /p), z = (sum(df_b[4] /p))/(sum(df_b[3]/p))))

@@

@@

```{r}
cat("Mean of the means of estimations obtained through each bootstrap process:\n - for the number of trees:", mean(EST_nb), "\n - for the volume of trees:", mean(EST_vol), "\n - for the ratio volume / number of trees:", mean(EST_ratio))
```

@@

```{r}
cat("Mean of the variance of bootstrap samples:\n - for the number of trees:", mean(VAR_nb), "\n - for the volume of trees:", mean(VAR_vol), "\n - for the ratio volume / number of trees:", mean(VAR_ratio))
```

@@

```{r}
diff_percent <- function(x,y)100*(max(x,y)-min(x,y))/min(x,y)
cat("Difference between the two estimations of variance: \n - for the number of trees: ", diff_percent(240516.7, mean(VAR_nb)), "%", "\n - for the volume of trees: ", diff_percent(14084.96, mean(VAR_vol)), "%\n - for the ratio volume / number of trees: ", diff_percent(6.397115e-06, mean(VAR_ratio)), "%", sep = "")
```
